---
title: "Credit Risk Scorecard Model - Model Development"
author: "Mayank Lal"
date: today
format: 
    html:
        code-fold: true
        code-summary: "Code"
        code-overflow: wrap
execute:
  echo: false
  include: false
  warning: false
---
```{r setup}
## Exploratory Data Analysis
library(tidyverse) # Data Manipulation and Visualizaton
library(data.table) # Large datasets
library(vtable) # Formatted Summary Table
library(corrplot) # Correlation
library(patchwork) # Grid of plots
library(scales) # For Percentage Labels
library(rpart) # Simple Decision Tree for interaction analysis
library(rpart.plot) # Plotting Decision Trees

## Variable Selection and Feature Engineering
library(scorecard) # Information Value
library(pROC) # Gini
library(gt) # Formattting Table for publishing
library(knitr) # For simple formatting
library(DT) # For interactive tables
library(stringr) # Selecting sub-strings
library(Hmisc) # Variabel Clustering
library(car) # VIF
library(glmnet) # LASSO

# Model Development
library(MASS) # Stepwise Regression
library(lmtest) # Joint Wald Test
library(Matrix) # Create Matrix objects for XGBoost
library(xgboost) # XGBoost
library(caret) # Stratified Splitting
library(lightgbm) # LightGBM

# Tidy Modeling
library(tidymodels)
library(recipes)
library(moments)
library(rlang)
library(vip)
library(future)
library(future.apply)
library(fastshap)
```

This is a rough sketch notebook. We will furnish it before publishing.
```{r setup}
set.seed(123)
```

```{r data}
# Load Model Data for Logistic Regression
load(file = "/workspace/Projects/kaggle-credit-risk-s5e11/data/processed/train_final_lr.Rdata")

# Load Test Data for Logistic Regression
load(file = "/workspace/Projects/kaggle-credit-risk-s5e11/data/processed/test_final_lr.Rdata")

# Load Model Data for Tree-based Models
load(file = "/workspace/Projects/kaggle-credit-risk-s5e11/data/processed/train_tree.Rdata")

# Load Test Data for Tree-based Models
load(file = "/workspace/Projects/kaggle-credit-risk-s5e11/data/processed/test_tree.Rdata")
```

## Model Development

Let's start with model development. First, we will use stepwise logistic regression to select final list of variables that improves AIC.

### Logistic Regression - Stepwise Regression (Iteration 0)

```{r lr_step}
# Create list of variables
vars <- c("working_flag_woe",
          "debt_to_income_ratio_woe",
          "credit_score_woe",
          "interest_rate_woe",
          "education_level_woe",
          "loan_purpose_woe",
          "annual_income_woe")

# Full model formula
full_formula <- as.formula(
  paste("loan_paid_back ~", paste(vars, collapse = " + "))
)

# Null model (only intercept)
null_formula <- loan_paid_back ~ 1

stepwise_model <- stepAIC(
    glm(loan_paid_back ~ ., data = train_final_lr, family = binomial),
    scope=list(lower=null_formula, upper=full_formula),
    direction="both",
    trace=TRUE
)

summary(stepwise_model)
```

### Logistic Regression - Final Model (Iteration 1)

```{r lr_final}
# Create list of final variables after stepwise regression
final_model_vars <- names(coef(stepwise_model))[-1]

# Final model formula
final_formula <- as.formula(
  paste("loan_paid_back ~", paste(final_model_vars, collapse = " + "))
)

# Build Final Logistic Regression Model
final_model <- glm(final_formula, data = train_final_lr, family = binomial)
summary(final_model)
```


#### Perform wald test for individual predictors

```{r wald_indv}
Anova(final_model)
```

### Logistic Regression - Final Model with interactions (Iteration 2)
```{r lr_interaction}
# Create interaction terms
interaction_terms <- c("debt_to_income_ratio_woe:credit_score_woe", "credit_score_woe:interest_rate_woe")

# Create final formula with interactions
final_formula_interaction <- as.formula(
  paste("loan_paid_back ~",
        paste(final_model_vars, collapse = " + "),
        "+",
        paste(interaction_terms, collapse = " + "))
)

# Build final model with interactions
final_model_interaction <- glm(final_formula_interaction,
                               data=train_final_lr,
                               family=binomial)
summary(final_model_interaction)
```

#### Perform wald joint test

```{r wald_joint}
waldtest(final_model, final_model_interaction)
```

### XGBoost - Base Model using k-fold cross-validation (Iteration 1)

In base model, we will be adding all variables in their raw form without any feature engineering.

Let's see the AUC of this base model. 

```{r xgb_cv_iter01}
# Create train and validation split 
train_idx <- createDataPartition(train_tree$loan_paid_back, p = 0.8, list = FALSE)
train_xgb <- train_tree[train_idx]
valid_xgb <- train_tree[-train_idx]

# Full formula 
feature_formula <- ~ . - id - loan_paid_back - 1

# Convert to DMatrix 
train_xgb_matrix <- model.matrix(feature_formula, data = train_xgb)
valid_xgb_matrix <- model.matrix(feature_formula, data = valid_xgb)

dtrain <- xgb.DMatrix(train_xgb_matrix, label = train_xgb$loan_paid_back)
dvalid <- xgb.DMatrix(valid_xgb_matrix, label = valid_xgb$loan_paid_back)

# Estimate Weight 
neg <- sum(train_xgb$loan_paid_back == 0)
pos <- sum(train_xgb$loan_paid_back == 1)
scale_pos_weight <- neg/pos

# Baseline parameters
params <- list(
  booster = "gbtree",
  objective = "binary:logistic",
  eval_metric = "auc",
  eta = 0.03,
  max_depth = 5,
  min_child_weight = 30,
  subsample = 0.8,
  colsample_bytree = 0.6,
  gamma = 1,
  lambda = 2,
  scale_pos_weight = scale_pos_weight
)

# Perform k-fold Cross-validation
xgb_cv <- xgb.cv(
  data = dtrain,
  param = params,
  nrounds = 5000,
  nfold = 10,
  early_stopping_rounds = 50,
  maximize = TRUE,
  verbose = 1
)

best_nrounds <- xgb_cv$early_stop$best_iteration

# Train Base Model
xgb_base_model <- xgb.train(
  params = params,
  data = dtrain,
  nrounds = best_nrounds,
  evals = list(train=dtrain, valid=dvalid),
  early_stopping_rounds = 50
)

# Predict on test dataset
feature_formula_test <- ~ . - id - 1
test_xgb_matrix <- model.matrix(feature_formula_test, data = test_tree)
test_pred_xgb  <- predict(xgb_base_model, test_xgb_matrix)
submission_xgb_iter01 <- data.frame(
  id = test_tree$id,
  loan_paid_back = test_pred_xgb
)

# Save to csv
write.csv(submission_xgb_iter01, "/workspace/Projects/kaggle-credit-risk-s5e11/outputs/submission_xgb_iter01.csv", row.names = FALSE)

# Feature Importance 
importance_matrix <- xgb.importance(
  model = xgb_base_model,
  feature_names = colnames(dtrain)
)

xgb.plot.importance(importance_matrix)
```

### XGBoost - Tidymodels (Iteration 2)

Based on discussions on Kaggle, I realized that my model development workflow is heavily flawed. <Will elaborate on why is it flawed later on.> To bridge the gap between my workflow and my peers' workflow, I will now use tidymodels based workflow. <Will elaborte on what is tidymodels and how it helps later on.>

```{r xgb_cv_iter02}
# Prepare datasets
vb_df <- train_tree %>% dplyr::select(-id)
vb_df$loan_paid_back <- factor(vb_df$loan_paid_back, levels = c(0, 1))
vb_split <- initial_split(vb_df, strata = loan_paid_back)
vb_train <- training(vb_split)
vb_test <- testing(vb_split)

# Specify paramters of XGBoost Model that needs to be tuned
xgb_spec <- boost_tree(
  trees = 1000,
  tree_depth = tune(), min_n = tune(),
  loss_reduction = tune(),                     ## first three: model complexity
  sample_size = tune(), mtry = tune(),         ## randomness
  learn_rate = tune()                          ## step size
) %>%
  set_engine("xgboost", eval_metric = "auc", device = "cuda") %>%
  set_mode("classification")

xgb_spec

# Specify hyperparameter space to find values for hyperparameters
xgb_grid <- grid_space_filling(
  tree_depth(),
  min_n(),
  loss_reduction(),
  sample_size = sample_prop(),
  finalize(mtry(), vb_train),
  learn_rate(),
  size = 30
)

xgb_grid

# Specify Workflow
xgb_rec <- recipe(loan_paid_back ~ ., data = vb_train) %>%
  step_zv(all_predictors()) %>%
  step_dummy(all_nominal_predictors())

xgb_wf <- workflow() %>%
  add_model(xgb_spec) %>%
  add_recipe(xgb_rec)


xgb_wf

# Specify folds for k-fold cross-validation
set.seed(123)
vb_folds <- vfold_cv(vb_train, strata = loan_paid_back, v = 3)

vb_folds

# Perform Hyperparameter tuning
doParallel::registerDoParallel()

set.seed(234)
xgb_res <- tune_grid(
  xgb_wf,
  resamples = vb_folds,
  grid = xgb_grid,
  control = control_grid(save_pred = TRUE)
)

xgb_res

# Explore metrics
collect_metrics(xgb_res)

# Plot metrics
xgb_res %>%
  collect_metrics() %>%
  filter(.metric == "roc_auc") %>%
  dplyr::select(mean, mtry:sample_size) %>%
  pivot_longer(mtry:sample_size,
               values_to = "value",
               names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "AUC")

# Find and save best metrics
show_best(xgb_res, metric = "roc_auc")
best_auc <- select_best(xgb_res, metric = "roc_auc")
best_auc

# Finalize tuneable workflow
final_xgb <- finalize_workflow(
  xgb_wf,
  best_auc
)

final_xgb

# Find Parameters 
final_xgb %>%
    fit (data = vb_train) %>%
    pull_workflow_fit() %>%
    vip(geom = "point")

# Evaluate model using best fit
final_res <- last_fit(final_xgb, vb_split)
collect_metrics(final_res)

# Create ROC curve
final_res %>%
    collect_predictions() %>%
    roc_curve(truth = loan_paid_back,.pred_1, event_level = "second") %>%
    ggplot(aes(x = 1 - specificity, y = sensitivity)) +
    geom_line(size = 1.5, color = "midnightblue") +
    geom_abline(
      lty = 2, alpha = 0.5,
      color = "gray50",
      size = 1.2
    )

# Extract workflow
final_wf <- final_res %>% extract_workflow()

# Predict on test dataset
test_pred <- predict(final_wf, test_tree, type = "prob")
submission_xgb_iter02 <- data.frame(
  id = test_tree$id,
  loan_paid_back = test_pred$.pred_1
)

# Save to csv
write.csv(submission_xgb_iter02, "/workspace/Projects/kaggle-credit-risk-s5e11/outputs/submission_xgb_iter02.csv", row.names = FALSE)
```


### XGBoost - Folds ensembled based Modeling using Target Encoding (Iteration 3)

This is my first attempt at advanced feature engineering as demonstrated by **Chris Deotte** in his [notebook](https://www.kaggle.com/code/cdeotte/feature-engineering-with-rapids-lb-38-847).

Earlier I tried to implement it separately and outside of my model training workflow. But I got limited success in that. So, I will be trying this again now within my training workflow using tidymodels.

Let's see how much success we get at this attempt. I have created following helper functions to perform advanced feature engineering:


- **Target encoding**: Encoding of categorical variables based on target variable

This code is written mostly by **ChatGPT** with bug fixes by me.

```{r xgb_cv_iter03}
# Load datasets
load(file = "/workspace/Projects/kaggle-credit-risk-s5e11/data/processed/train_fe.Rdata")
load(file = "/workspace/Projects/kaggle-credit-risk-s5e11/data/processed/test_fe.Rdata")

dt_train <- setDT(train_fe)
dt_test <- setDT(test_fe)

# Load target encoding function
source("/workspace/Projects/kaggle-credit-risk-s5e11/helpR_funcs/fe_target_encoder.R")

# Load ensemble fold cv model function
source("/workspace/Projects/kaggle-credit-risk-s5e11/helpR_funcs/ensemble_fold_cv_model.R")

# Define target variable 
target <- "loan_paid_back"

# Define id variable
id <- "id"

# Define non-features
non_features <- list(id, target)

# Categorise features
features <- names(dt_train[, .SD, .SDcols = setdiff(names(dt_train), non_features)])
numeric <- names(dt_train[, .SD, .SDcols = setdiff(names(dt_train[,.SD,.SDcol = is.numeric]), non_features)])
factors <- names(dt_train[,.SD, .SDcols = setdiff(names(dt_train), union(numeric, union(target, non_features)))])
inter <- colnames(dt_train %>% dplyr::select(starts_with("inter_")))
round <- colnames(dt_train %>% dplyr::select(starts_with("round_")))

# Specify XGBoost Params
xgb_params <- list(
  booster = "gbtree",
  
  # === Objective ===
  objective = "binary:logistic",
  eval_metric = "auc",
  
  # === Main hyperparameters ===
  eta = 0.01,               # learn_rate
  max_depth = 6,            # tree_depth
  subsample = 0.55,         # sample_size
  colsample_bytree = 0.3,   # mtry equivalent
  gamma = 0,                # loss_reduction
  
  # === GPU (CUDA) settings ===
  device = "cuda",
  
  # === System settings ===
  nthread = parallel::detectCores()
)

# Perform Ensemble Fold CV Model Run 
res <- ensemble_fold_cv_model(
  train_dt = dt_train,
  test_dt  = dt_test,
  target   = target,

  INTER = inter,
  ROUND = round,
  CATS  = factors,

  xgb_params = xgb_params,
  folds = 5,
  te_cv = 5,
  te_smooth = 1.0,
  drop_original_te1 = TRUE,
  drop_original_te2 = FALSE,

  parallel = FALSE,
  seed = 2025,
  verbose = TRUE
)

# Extract submissions
submission_xgb_iter03 <- data.frame(
  id = dt_test$id,
  loan_paid_back = res$test_preds
)

# Save to csv
write.csv(submission_xgb_iter03, "/workspace/Projects/kaggle-credit-risk-s5e11/outputs/submission_xgb_iter03.csv", row.names = FALSE)
```


### XGBoost - Folds ensembled based Modeling using Target Encoding after pruning based on Gain and SHAP (Iteration 4)

In this iteration, we will apply pruning of tree to remove unwanted features using SHAP and retrain our model.

```{r shap_pruning}
# Build fully transformed train dataset from previous iteration
dtrain_full <- xgb.DMatrix(data = res$X_full, label = res$y_full)

# Train Full Model before pruning
xgb_pre_prune_model <- xgb.train(
  params = xgb_params,
  data = dtrain_full,
  nrounds = 1500
)

# Compute XGB Importance
importance_matrix <- xgb.importance(
  model = xgb_pre_prune_model,
  feature_names = colnames(dtrain_full)
)

xgb.plot.importance(importance_matrix[1:20,])

# Store Gain scores
gain_scores <- setNames(importance_matrix$Gain, importance_matrix$Feature)

gain_importance <- data.table(
  Feature = names(gain_scores),
  Gain = as.numeric(gain_scores)
)

# Construct VIP using "exact" SHAP values from XGBoost's internal Tree SHAP
# functionality

vip(
  xgb_pre_prune_model,
  method = "shap",
  train = res$X_full,
  exact = TRUE,
  num_features = 20,
  include_type = TRUE,
  geom = "point",
  horizontal = TRUE
) + theme_light()

vip_shap_tbl <- vip::vi(
  object = xgb_pre_prune_model,
  method = "shap",
  train = res$X_full,
  exact = TRUE
)

shap_importance <- data.table(
  Feature = vip_shap_tbl$Variable,
  SHAP    = vip_shap_tbl$Importance
)

# Combine Gain and SHAP into importance table 
importance <- merge(
  shap_importance,
  gain_importance,
  by = "Feature",
  all = TRUE
)

# Replace NA with 0
importance[is.na(importance)] <- 0

importance[, Combined := 0.6*Gain + 0.4*SHAP]
importance <- importance[order(-Combined)]

# Select features to drop
importance_pruned <- importance[Combined >= 0.01]
keep_features <- importance_pruned$Feature
drop_features <- importance[Combined < 0.01]$Feature

# Clean feature names
clean_names <- gsub("^TE_", "", keep_features)
clean_features <- gsub("_mean$", "", clean_names)
```


```{r xgb_cv_iter04}
# Clean datasets 
dt_train_clean <- dt_train[, .SD, .SDcols = union(clean_features, target)]
dt_test_clean <- dt_test[, .SD, .SDcols = clean_features]

# Categorise features
features <- clean_features
numeric <- names(dt_train_clean[, .SD, .SDcols = setdiff(names(dt_train_clean[,.SD,.SDcol = is.numeric]), non_features)])
factors <- names(dt_train_clean[,.SD, .SDcols = setdiff(names(dt_train_clean), union(numeric, union(target, non_features)))])
inter <- colnames(dt_train_clean %>% dplyr::select(starts_with("inter_")))
round <- colnames(dt_train_clean %>% dplyr::select(starts_with("round_")))

# Specify XGBoost Params
xgb_params <- list(
  booster = "gbtree",
  
  # === Objective ===
  objective = "binary:logistic",
  eval_metric = "auc",
  
  # === Main hyperparameters ===
  eta = 0.01,               # learn_rate
  max_depth = 6,            # tree_depth
  subsample = 0.55,         # sample_size
  colsample_bytree = 0.3,   # mtry equivalent
  gamma = 0,                # loss_reduction
  
  # === GPU (CUDA) settings ===
  device = "cuda",
  
  # === System settings ===
  nthread = parallel::detectCores()
)

# Perform Ensemble Fold CV Model Run 
res_v4 <- ensemble_fold_cv_model(
  train_dt = dt_train_clean,
  test_dt  = dt_test_clean,
  target   = target,

  INTER = inter,
  ROUND = round,
  CATS  = factors,

  xgb_params = xgb_params,
  folds = 5,
  te_cv = 5,
  te_smooth = 1.0,
  drop_original_te1 = TRUE,
  drop_original_te2 = FALSE,

  parallel = FALSE,
  seed = 2025,
  verbose = TRUE
)

# Extract submissions
submission_xgb_iter04 <- data.frame(
  id = dt_test$id,
  loan_paid_back = res_v4$test_preds
)

# Save to csv
write.csv(submission_xgb_iter04, "/workspace/Projects/kaggle-credit-risk-s5e11/outputs/submission_xgb_iter04.csv", row.names = FALSE)

# Create full dataset
dtrain_full <- xgb.DMatrix(data = res$X_full, label = res$y_full)

# Train final XGBoost Model
final_xgb_model <- xgb.train(
  params = xgb_params,
  data = dtrain_full,
  nrounds =  1800, # or manual
  verbose = 1
)

# Save model
xgb.save(final_xgb_model, "models/xgb/xgb_final.json")
saveRDS(colnames(res$X_full), "models/xgb/xgb_feature_list_full.rds")
write.csv(res_v4$oof_preds, "models/xgb/OOF_xgb.csv", row.names = FALSE)
```

### LightGBM - Base Model with all features and more features based on quantile binning, high signal encoding and numeric interactions (Iteration 1)


```{r lgbm_iter01}
# Load datasets
load(file = "/workspace/Projects/kaggle-credit-risk-s5e11/data/processed/train_fe_v2.Rdata")
load(file = "/workspace/Projects/kaggle-credit-risk-s5e11/data/processed/test_fe_v2.Rdata")

dt_train_lgbm <- setDT(train_fe_v2)
dt_test_lgbm <- setDT(test_fe_v2)

# Load target encoding function
source("/workspace/Projects/kaggle-credit-risk-s5e11/helpR_funcs/fe_target_encoder.R")

# Load logistic regression encoding function
source("/workspace/Projects/kaggle-credit-risk-s5e11/helpR_funcs/logistic_regression_encode.R")

# Load ensemble fold cv model function
source("/workspace/Projects/kaggle-credit-risk-s5e11/helpR_funcs/ensemble_fold_cv_lgb.R")

# Define target variable 
target <- "loan_paid_back"

# Define id variable
id <- "id"

# Define non-features
non_features <- list(id, target)

# Remove ID column
dt_train_lgbm <- dt_train_lgbm[,.SD,.SDcols = -id]
dt_test_lgbm <- dt_test_lgbm[,.SD,.SDcols = -id]

# Full feature list
#full_features <- colnames(dt_train_lgbm)

# XGB Model features
#xgb_features <- clean_features

# Categorise features
factors_raw <- c("grade", "subgrade", "employment_status", "education_level", "loan_purpose", "education_level", "marital_status", "gender")
numeric <- names(dt_train_lgbm[, .SD, .SDcols = setdiff(names(dt_train_lgbm[,.SD,.SDcol = is.numeric]), non_features)])
inter <- colnames(dt_train_lgbm %>% dplyr::select(starts_with("inter_")))
round <- colnames(dt_train_lgbm %>% dplyr::select(starts_with("round_")))

# Specify LightGBM Params
lgb_params <- list(
  objective          = "binary",
  metric             = "auc",
  boosting_type      = "gbdt",
  
  learning_rate      = 0.01,
  num_leaves         = 31,
  max_depth          = -1,
  
  # Auto-pruning controls
  min_data_in_leaf   = 50,
  min_gain_to_split  = 0.01,
  reg_alpha          = 0.2,
  reg_lambda         = 0.5,
  
  # Sampling-based auto-pruning
  colsample_bytree   = 0.6,
  colsample_bynode   = 0.6,
  subsample          = 0.7,
  subsample_freq     = 1,
  
  # Speed / reproducibility
  num_threads        = parallel::detectCores(),
  device             = "gpu",
  verbose            = 1,
  seed               = 42
)


# ================================
# Perform Ensemble Fold CV Model Run (LightGBM)
# ================================

# Let's drop inter variables and then run it

res_lgb <- ensemble_fold_cv_lgb(
  train_dt = dt_train_lgbm,
  test_dt  = dt_test_lgbm,
  target   = target,

  # --- Feature groups ---
  RAW_CATS = factors_raw,     # raw categorical vars for LR embeddings
  INTER    = inter,           # inter_* for TE (if enabled)
  ROUND    = round,           # round_* for TE (if enabled)
  enable_te = TRUE,           # switch TE on/off for testing performance impact

  # --- LightGBM parameters (your tuned list) ---
  lgb_params = lgb_params,    # <--- the param list you prepared earlier

  # --- CV settings ---
  folds = 5,
  te_cv = 5,
  te_smooth = "auto",
  lambda_lr = 1e-6,           # LR embedding ridge penalty
  drop_original_te1 = TRUE,
  drop_original_te2 = FALSE,

  # --- Runtime ---
  parallel = FALSE,           # or TRUE if needed
  seed = 2025,
  verbose = TRUE,

  # --- Training loop ---
  nrounds_local = 10000,       # booster rounds (equivalent to num_boost_round)
  early_stopping_rounds = 100, # stop like Python's early_stopping(100)

  # --- Debug / Model Saving (optional) ---
  debug = TRUE,
  snapshot_dir = "models/lgbm",
  model_dir = "models/lgbm"     # saves each fold model
)

# Build fully transformed train dataset from previous iteration
dtrain <- lightgbm::lgb.Dataset(data = res_lgb$X_full, label = res_lgb$y_full)
dval <- lightgbm::lgb.Dataset(data = res_lgb$X_full, label = res_lgb$y_full, reference = dtrain)

# Train Full Model before pruning
lgb_pre_prune_model <- suppressWarnings(
                        lightgbm::lgb.train(
                          params = lgb_params,
                          data = dtrain,
                          nrounds = 1216,
                          verbose = 1L
                        )
                      )

# -----------------------------------------------------------
# 1. LightGBM GAIN Importance
# -----------------------------------------------------------

lgb_imp <- lightgbm::lgb.importance(lgb_pre_prune_model, percentage = TRUE)

gain_importance <- data.table(
  Feature = lgb_imp$Feature,
  Gain = as.numeric(lgb_imp$Gain)
)

# -----------------------------------------------------------
# 2. SHAP Importance via vip::vi()
#    vip supports LightGBM models directly.
# -----------------------------------------------------------

vip_shap_tbl <- vip::vi(
  object = lgb_pre_prune_model,
  method = "shap",
  train  = res_lgb$X_full,
  exact  = TRUE,
  feature_names = colnames(res_lgb$X_full)
)

shap_importance <- data.table(
  Feature = vip_shap_tbl$Variable,
  SHAP    = vip_shap_tbl$Importance
)

# -----------------------------------------------------------
# 3. Merge SHAP + Gain
# -----------------------------------------------------------

importance <- merge(
  shap_importance,
  gain_importance,
  by = "Feature",
  all = TRUE
)

importance[is.na(importance)] <- 0

# Weighted combined importance (same as your XGB logic)
importance[, Combined := 0.6 * Gain + 0.4 * SHAP]

importance <- importance[order(-Combined)]

# -----------------------------------------------------------
# 4. Threshold for pruning
# -----------------------------------------------------------

threshold <- 0.01  # tweak if needed

importance_pruned <- importance[Combined >= threshold]

keep_features <- importance_pruned$Feature
drop_features <- importance[Combined < threshold]$Feature

# -----------------------------------------------------------
# 5. Output Summary
# -----------------------------------------------------------

cat("Total features:", nrow(importance), "\n")
cat("Keeping:", length(keep_features), "\n")
cat("Dropping:", length(drop_features), "\n")

head(importance)
```

```{r lgbm_iter_02}
# Clean feature names
# Identify only TE_ columns
te_cols <- grep("^TE_", keep_features, value = TRUE)
te_cols_cleaned_v1 <- gsub("^TE_", "", te_cols)
te_cols_cleaned <- gsub("_mean$", "", te_cols_cleaned_v1)
cols_not_te <- setdiff(keep_features, te_cols)
cols_not_te_clened <- gsub("^lr_cat_", "", cols_not_te)

clean_features <- union(te_cols_cleaned, cols_not_te_clened)

# Clean datasets 
dt_train_lgbm_clean <- dt_train_lgbm[, .SD, .SDcols = union(clean_features, target)]
dt_test_lgbm_clean <- dt_train_lgbm[, .SD, .SDcols = clean_features]

# Categorise features
factors_raw <- "employment_status"
numeric <- names(dt_train_lgbm_clean[, .SD, .SDcols = setdiff(names(dt_train_lgbm_clean[,.SD,.SDcol = is.numeric]), non_features)])
inter <- colnames(dt_train_lgbm_clean %>% dplyr::select(starts_with("inter_")))
round <- colnames(dt_train_lgbm_clean %>% dplyr::select(starts_with("round_")))


# ================================
# Perform Ensemble Fold CV Model Run (LightGBM)
# ================================

# Let's drop inter variables and then run it

res_lgb_v2 <- ensemble_fold_cv_lgb(
  train_dt = dt_train_lgbm_clean,
  test_dt  = dt_test_lgbm_clean,
  target   = target,

  # --- Feature groups ---
  RAW_CATS = factors_raw,     # raw categorical vars for LR embeddings
  INTER    = inter,           # inter_* for TE (if enabled)
  ROUND    = round,           # round_* for TE (if enabled)
  enable_te = TRUE,           # switch TE on/off for testing performance impact

  # --- LightGBM parameters (your tuned list) ---
  lgb_params = lgb_params,    # <--- the param list you prepared earlier

  # --- CV settings ---
  folds = 5,
  te_cv = 5,
  te_smooth = "auto",
  lambda_lr = 1e-6,           # LR embedding ridge penalty
  drop_original_te1 = TRUE,
  drop_original_te2 = FALSE,

  # --- Runtime ---
  parallel = FALSE,           # or TRUE if needed
  seed = 2025,
  verbose = TRUE,

  # --- Training loop ---
  nrounds_local = 10000,       # booster rounds (equivalent to num_boost_round)
  early_stopping_rounds = 100, # stop like Python's early_stopping(100)

  # --- Debug / Model Saving (optional) ---
  debug = TRUE,
  snapshot_dir = "models/lgbm",
  model_dir = "models/lgbm"     # saves each fold model
)

```